Convolutional neural nets for the MNIST digit dataset

### 2014-12-12 tex/paper.tex
Reorg and small paper updates.
Reorganized the folder layout because there were
just too many files on my machine.
Made small edits with this reorg: read_mnist points to
a different place, removed commented out code from MATLAB visualization
code, and added trained config files for oneConv.
Also added a new subsection CNN's to section 2 of the paper and
added a small amount of new language to the FC description in section 3.

%%%%%%%%% TITLE
\title{New titles are hard to come up with}

@@

@@

To address the scalability and translational invariance issues inherent in DBNs,
Lee \etal introduce the convolutional deep belief network (CDBN),
which augments DBNs with convolutional restricted Boltzmann machines (CRBMs)
and probabilistic max-pooling layers between CRBMs \cite{CDBN}.
Like RBMs, CRBMs consist of a visible layer and hidden layer, but the hidden
layer here is partitioned into $K$ square blocks, each associated with a filter
whose weights are shared amongst the units in a block.
The relation between the visible units and the corresponding hidden units in each
block is defined by a convolution of the weight matrix and the image.

To alleviate the scaling issue, a probabilistic max-pooling layer is added above the CRBM,
forming a max-pooling CRBM.
This shrinks the representation of the hidden layer by a constant factor.
These pooling layers contain $K$ blocks, with each block mapping to the corresponding hidden
layer block below it as follows: the hidden layer block is partitioned into square blocks with the
corresponding pooling unit for each block being set to 1 if and only if a unit in its hidden layer
block is 1, of which only one can be non-zero.
The CDBN is constructed by stacking max-pooling CRBMs,
which are trained similarly to RBMs.

@@

Because CDBNs construct a hierarchical representation of their input that is robust
to appearance, scale, and occlusion variation, they are are able to deliver
state-of-the-art performance.
They are not without shortcomings: in particular, they suffer from over-fitting due
to over-representation of their input data.
There are several solutions to this problem. In dropout, the output of a hidden layer
unit is set to 0 with probability $\frac{1}{2}$; this unfortunately significantly increases training
time (which can already take weeks). Another common solution, data augmentation,
is to insert altered forms of the input images into the dataset (used in \cite{Verydeep}).

Modifications to the original CDBN architecture \cite{CDBN} have produced the most performant
systems to date.
By overlapping hidden layer blocks during max-pooling, \cite{ImageNet} increases performance.
The approach outlined in \cite{Verydeep} gains significant accuracy by adding many extra layers
to the CDBN and emphasizing very small filters.

In order to counter the extra complexity introduced by these changes,
\cite{ImageNet} and \cite{Verydeep} replace the standard sigmoid nonlinear output function with the
Rectified Linear Unit (ReLU) function $f(x) = max(0, x)$ for quicker training.
Additionally, both use multiple GPUs for significant speedup in matrix operations
and easy parallelism.
Nevertheless, these CDBNs take a very long time to train: the net in \cite{Verydeep}
takes 2-3 weeks to train using four GPUs.

@@

@@

The following experiments were conducted using the MNIST dataset [?], which includes a training set of 
60000 28x28 handwritten digit images and a test set of 10000 28x28 handwritten digit images. It is important 
to note that both the training and test sets are composed of two sets of digits: an easier set written by 
US Census Bureau employees and a harder set written by high school students.

>>>>

%%%%%%%%% TITLE
\title{Convolutional neural nets for the MNIST digit dataset}

@@

%-------------------------------------------------------------------------
\subsection{Convolutional Neural Networks}

We can construct a different way to use a hierarchical representation of an image for classification
by performing convolutions at different scales of images with sets of filters.
Convolutional Neural Nets (CNNs) use a combination of convolutional layers, which perform convolution
of an image with a set of filters, and pooling layers, which reduce the dimensionality of the output
of convolutional layers, to create this representation.
Pooling layers reduce dimensionality by taking the maximum value (max-pooling) or mean value
(mean-pooling) of a region of convolutional outputs.
This reduction of dimensionality is important because it helps reduce the number of parameters that
must be trained in the network, thus speeding up learning.

Modifications to this architecture \cite{CDBN} have produced some of the most performant
systems to date.
By overlapping hidden layer blocks during max-pooling, \cite{ImageNet} increases performance.
The approach outlined in \cite{Verydeep} gains significant accuracy by constructing a net with
a large number of convolution and pooling layers (using very small filters).

In order to counter the extra complexity introduced by these changes,
\cite{ImageNet} and \cite{Verydeep} replace the standard sigmoid nonlinear output function with the
Rectified Linear Unit (ReLU) function $f(x) = max(0, x)$ for quicker training.
Additionally, both use multiple GPUs for significant speedup in matrix operations
and easy parallelism.
Nevertheless, these CNNs take a very long time to train: the net in \cite{Verydeep}
takes 2-3 weeks to train using four GPUs.

@@

The idea of using convolution can also be applied to Deep Belief nets.
To address the scalability and translational invariance issues inherent in DBNs,
Lee \etal introduce the convolutional deep belief network (CDBN),
which augments DBNs with convolutional restricted Boltzmann machines (CRBMs)
and probabilistic max-pooling layers between CRBMs \cite{CDBN}.
Like RBMs, CRBMs consist of a visible layer and hidden layer, but the hidden
layer here is partitioned into $K$ square blocks, each associated with a filter
whose weights are shared amongst the units in a block.
The relation between the visible units and the corresponding hidden units in each
block is defined by a convolution of the weight matrix and the image.

To alleviate the scaling issue, a probabilistic max-pooling layer is added above the CRBM,
forming a max-pooling CRBM.
These pooling layers contain $K$ blocks, with each block mapping to the corresponding hidden
layer block below it as follows: the hidden layer block is partitioned into square blocks with the
corresponding pooling unit for each block being set to 1 if and only if a unit in its hidden layer
block is 1, of which only one can be non-zero.
The CDBN is constructed by stacking max-pooling CRBMs, which are trained similarly to RBMs.

@@

Because CDBNs construct a hierarchical representation of their input that is robust
to appearance, scale, and occlusion variation, they are are able to deliver
state-of-the-art performance.
They are not without shortcomings: in particular, they suffer from over-fitting due
to over-representation of their input data.
There are several solutions to this problem. In dropout, the output of a hidden layer
unit is set to 0 with probability $\frac{1}{2}$; this unfortunately significantly increases training
time (which can already take weeks). Another common solution, data augmentation,
is to insert altered forms of the input images into the dataset
(a good description of this idea is provided in \cite{Verydeep}).

@@

Because a single input image is broken up into multiple samples during convolution, our
fully connected layer is capable of batch training, where the update to a set of weights over
many samples is the sum of updates for each sample. While batch training as a whole is generally
slower than on-line training \cite{EfficientBackProp}, we are still performing on-line training
with respect to each input image, so we do not take an unreasonable performance hit.

@@

The following experiments were conducted using the MNIST dataset \cite{MNIST}, which includes a training set of
60000 28x28 handwritten digit images and a test set of 10000 28x28 handwritten digit images. It is important
to note that both the training and test sets are composed of two sets of digits: an easier set written by
US Census Bureau employees and a harder set written by high school students.
